### 2020-02-08 thesis/paper.tex
more

The meta-view consists of five visual elements and some UI-elements in the top. Of the five elements, the left four visualizations give an overview of the data on an individual clustering and meta-level and the right part can be used to filter the base clusterings. In the top left visualization, the data can be seen with the clustering labels colored in according to the result of selected clustering run. The selected clustering can be changed by either selecting another one in the selector above, by clicking on a result in one of the other three views that  show them on a meta-level. When changing between selected clustering solutions, the colors for the labels are adapted such that the change of labels is minimal. This is done so the user can more easily see differences, and is computed solving the mathematical assignment problem with Kuhn’s Hungarian method \cite{Kuhn2010}.

@@

\begin{itemize}
  \item Clustering Distance Measure: IMetaDistanceMeasure
\end{itemize}
\subsubsection{Clustering Distance Measures}
The interface IMetaDistanceMeasure is used to define distance measures that can be used for OPTICS for mata-clustering. These distance measures should represent how different two clustering solutions are from each other while needing to fulfill the requirements of a metric. For a distance measure, here indicated with $d(x,y)$ where $x$ and $y$ are objects, to be considered metric the following conditions must hold:
\begin{itemize}
  \item $d(x,y)\geq0$: distances must be non-negative
  \item $d(x,y)=0$ iff $x=y$: the distance between two objects is zero if and only if both objects are equal
  \item $d(x,y)=d(x,y)$: the distance must fulfill the symmetry condition
  \item $d(x,z)\leq d(x,y)+d(y,z)$: the triangle inequality must hold
\end{itemize}

@@

\subsection{OPTICS Plot}

bla bla bla

\subsection{Heat Map}

bla bla bla

@@

\subsection{Filter Panel}

bla bla bla

>>>>

The meta-view consists of five visual elements and some UI-elements in the top. Of the five elements, the left four visualizations give an overview of the data on an individual clustering and meta-level and the right part can be used to filter the base clusterings. In the top left visualization, the data can be seen with the clustering labels colored in according to the result of selected clustering run. The selected clustering can be changed by either selecting another one in the selector above, by clicking on a result in one of the other three views that  show them on a meta-level. When changing between selected clustering solutions, the colors for the labels are adapted such that the change of labels is minimal. This is done so the user can more easily see differences, and is computed solving the mathematical assignment problem with Kuhn’s Hungarian method \cite{Kuhn2010}. If the user wishes to reset the colors for the labels of the currently selected coloring, for example when the recoloring leads to similar colors, he can press the ``C'' button in the top right of the view, setting the colors to their default values.

@@

\begin{itemize}
  \item Clustering Distance Measure: IMetaDistanceMeasure
\end{itemize}
\subsubsection{Clustering Distance Measures}
The interface IMetaDistanceMeasure is used to define distance measures that can be used for OPTICS for mata-clustering. These distance measures should represent how different two clustering solutions are from each other while needing to fulfill the requirements of a metric. For a distance measure, here indicated with $d(x,y)$ where $x$ and $y$ are objects, to be considered metric the following conditions must hold \cite{10.5555/1756006.1953024}:
\begin{itemize}
  \item $d(x,y)\geq0$: distances must be non-negative
  \item $d(x,y)=0$ iff $x=y$: the distance between two objects is zero if and only if both objects are equal
  \item $d(x,y)=d(x,y)$: the distance must fulfill the symmetry condition
  \item $d(x,z)\leq d(x,y)+d(y,z)$: the triangle inequality must hold
\end{itemize}

@@

\subsection{OPTICS Plot}
The OPTICS reachability plot shows the clusterings with their computed reachability values computed by the OPTICS algorithm ordered according to their order in which they were processed. For additional information on OPTICS see Section \ref{sec:opticstheory}. Clicking or dragging on the gray range bar changes the threshold for which clusterings are considered to be in a meta-cluster and the colors are adapted accordingly. The colors of the clusterings when represented in the MDS plot are also defined through this threshold. The OPTICS plot, just like the heat map and the MDS plot allow for selecting clusterings through clicking or control-clicking on them, whenever multiple clusterings should be selected.

The implementation of the OPTICS algorithm, creating the data required for this visualization is based on the proposal of Ankerst et al., using a priority queue to efficiently keep a fitting order for the clusterings that need to be processed.

\subsection{Heat Map}
The heat map shows the distances between clusterings solutions using the meta distance measure as defined by the user. It allows for visually identifying blocks of similar results on the diagonal, different hues of brown were chosen for this as they are easier to distinguish than other colors tested. The plot is drawn by overriding Java Swings \cite{javaswing} draw() function as creating a component for each cell makes the view unresponsive due to the large overhead from managing a large number of components. To allow for mouse interactivity the coordinates of clicks were calculated back to find the targeted cell and used for selection logic. This is needed as there are no small components which would usually manage this.

@@

\subsection{Filter Panel}
The filter panel allows the user to filter out results by decreasing the range on the range selectors. The area charts show how much data points are within the range in blue and how many points are outside it/filtered out in red. This chart is implemented by defining buckets for the values to be shown and counting the number of value occurrences within each bucket. For double values the square root of the number of values is used as the number of buckets, for integer values the number of values in their range is chosen and for Boolean values 3 buckets are chosen. 3 buckets where chosen there, such that true is shown on the left, false on the right and there is space in the center for clearly separating the two values. If there would be to many buckets when using this default calculation they are capped at 31 (as uneven numbers may produce nicer ranges), which helps for example whenever too large integer ranges are available. Also, as the charts are quite small, more buckets can no longer be visualized properly.

For each clustering algorithm type a separate chart and range selector is created for each parameter and some additional filters are available. The parameters are found through the Parameter object which must be added to a clustering result by the class performing the clustering algorithm. The additional filters include the distance (as chosen by the user) to the ground truth and the Normalized Mutual Information (NMI) \cite{10.5555/1756006.1953024} whenever ground truth is available, as well as the ratio of the size of the largest cluster relative to the number of points. This ratio may be useful as clusterings where e.g. if more than $90\%$ of the points are within one cluster, the solution may not be satisfactory. More such filters can be added by adding the needed values to the clustering results through the addParameters() function in the MetaView class.

When the range is reduced, filtering out results, the other views are updated to reflect this change by changing the opacity of the corresponding points/cells.
